---
title: "Penguin"
author: "Josh Kong"
date: "12/2/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
# Objective: Create a classification model that predicts the sex of a penguin based on certain features.

# Loading the data and neccesary packages in
```{r}
penguins <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')
library(tidyverse)
library(tidymodels)
```

# Data exploration
Looking at the distribution of different penguin species
```{r}
penguins %>% 
  count(species) %>%
  mutate(species = fct_reorder(species, n)) %>% 
  ggplot(aes(species, n, fill = "blue")) + 
  geom_col() + 
  theme(legend.position = "none") +
  labs(x = "Species", y = "Count", title = "Distribution of Penguin Species")
```

Taking a look at the relationship between bill length and flipper length of penguins of different sex and different species
```{r}
penguins %>% 
  filter(!is.na(sex)) %>% 
  ggplot(aes(bill_length_mm, flipper_length_mm, color = sex, size = body_mass_g)) + 
  geom_point(alpha = 0.7) + 
  facet_wrap(~species) + 
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)", title = "Relationship Between Bill and Flipper Length of Penguins", color = "Sex", size = "Body Mass (g)") 
```

Selecting the data we want for our model
```{r}
penguins_df <- penguins %>% 
  filter(!is.na(sex)) %>% 
  select(-year, -island)

#changing our character columns into factors
penguins_df <- penguins_df %>% 
  mutate_if(is.character, factor)
```



# Model Building

## Splitting the data
```{r}
set.seed(123)
penguin_split <- initial_split(penguins_df, strata = sex)
penguin_train <- training(penguin_split)
penguin_test <- testing(penguin_split)
```

Creating a recipe
```{r}
penguins_df
penguin_rec <- recipe(sex ~ ., data = penguin_train) %>% 
  step_dummy(species,one_hot = TRUE)   #making species into dummy variables.
```


## KNN Model

Building the Model
```{r}
knn_spec <- nearest_neighbor(neighbors = 9) %>% 
  set_mode("classification") %>% 
  set_engine("kknn")

knn_wf <- workflow() %>% 
  add_recipe(penguin_rec) %>% 
  add_model(knn_spec)

knn_fit <- knn_wf %>% 
  fit(penguin_train)
```

Evaluating the Model
```{r}
pred_knn <- predict(knn_fit, penguin_test)
knn_conf <- table(pred_knn$.pred_class, penguin_test$sex); knn_conf
knn_acc <- (knn_conf[1,1] + knn_conf[2,2]) / sum(knn_conf); knn_acc

paste0("Got an accuracy of ",round(knn_acc,2),"% using KNN.")
```



## Random Forest Model

Building the model
```{r}
set.seed(123)
rf_spec <- rand_forest(trees= 1000, mtry = 4) %>% 
  set_mode("classification") %>% 
  set_engine("ranger")

rf_wf <- workflow() %>% 
  add_recipe(penguin_rec) %>% 
  add_model(rf_spec)

rf_fit <- rf_wf %>% 
  fit(penguin_train)
```

Evaluating the Model
```{r}
pred_rf <- predict(rf_fit, penguin_test)
rf_conf <- table(pred_rf$.pred_class, penguin_test$sex); rf_conf
rf_acc <- (rf_conf[1,1] + rf_conf[2,2]) / sum(rf_conf); rf_acc

paste0("Got an accuracy of ",round(rf_acc,2),"% using a random forest model.")
```


## Boosted Tree Model

Building the Model
```{r}
set.seed(234)
xgb_spec <- boost_tree(trees= 1000, mtry = 7) %>% 
  set_mode("classification") %>% 
  set_engine("xgboost")

xgb_wf <- workflow() %>% 
  add_recipe(penguin_rec) %>% 
  add_model(xgb_spec)

xgb_fit <- xgb_wf %>% 
  fit(penguin_train)
```

Evaluating the Model
```{r}
pred_xgb <- predict(xgb_fit, penguin_test)
xgb_conf <- table(pred_xgb$.pred_class, penguin_test$sex); xgb_conf
xgb_acc <- (xgb_conf[1,1] + xgb_conf[2,2]) / sum(xgb_conf); xgb_acc

paste0("Got an accuracy of ",round(xgb_acc,2),"% using a boosted tree model.")
```



```{r}
#which variables contributed most to the classfication of the sex of penguins?
library(vip)
xgb_fit %>% 
  pull_workflow_fit() %>% 
  vip(geom="point")
```

The two most important variables in this classfication model were the bill depth of the penguin and the body mass of the penguin.


# Conclusion

I used 3 classification machine learning models (K-Nearest Neighbor, Random Forest Classifier, Boosted Tree Classifier) to predict the sex of a penguin based on certain features. Using a random forest classifier, I was able to predict the sex of a penguin with an accuracy rate of ~95%.

My final conclusion is, the feature that strongly defines a penguin's sex is the bill depth and the body mass. 

















